---
title: "earthtones"
output: github_document
---

```{r setup, include=FALSE}
if(!require(devtools)) install.packages("devtools")
install_github("wcornwell/earthtones")
library(earthtones)
```

## Geographic color schemes

Let's say you wanted a color scheme based on a particular part of the world.  For example the grand canyon.  

```{r grand canyon}
grand_canyon<-plot_satellite_image_and_pallette(latitude = 36.094994,longitude=-111.837962,zoom=12,number_of_colors=5)
```

Or maybe you want a color scheme drawn from tropical reefs and lagoons.  

```{r bahamas}
bahamas<-plot_satellite_image_and_pallette(latitude = 24.2,longitude=-77.88,zoom=11,number_of_colors=5)
```

Just pick your favorite place in the world, and find out the major colors

```{r uluru}
 uluru<-plot_satellite_image_and_pallette(latitude = -25.5,longitude = 131,zoom=10,number_of_colors=5)
```

The function `plot_satellite_image_and_pallette` is good for seeing both the image and the color palette.  To actually use the color, it's much easier to use `get_earthtones`.  For example:

```{r }
library(ggplot2)
bahamas_colors<-get_earthtones(latitude = 24.2,longitude=-77.88,
zoom=11,number_of_colors=3)
ggplot(iris,aes(x=Petal.Length,y=Petal.Width,col=Species))+geom_point(size = 2)+
  scale_color_manual(values = bahamas_colors)+theme_bw()
```

or

```{r}
grand_canyon<-get_earthtones(latitude = 36.094994,longitude=-111.837962,zoom=12,number_of_colors=3)
ggplot(iris,aes(x=Petal.Length,y=Petal.Width,col=Species))+geom_point(size = 2)+
  scale_color_manual(values = grand_canyon)+theme_bw()
```

##Methods details

1. This library gets an image from Google earth which come from different sources depending on the zoom and the particular place.  

2. It then extracts the colors in the image, translates them into a perceptually uniform color space and then runs k-means clustering algorithm to find the major colors for an area

3. These are then converted back into a R style color pallete.  